On the analytic structure of the linear response susceptibility in open systems

### 2018-04-26 AnalyticStructure_main.tex

We restrict our analysis assuming that a set of hypothesis commonly realized in DFT holds (symmetry of the transition basis, specify better? see appendix?). Equations 
\eqref{ExcitationOperatorsDef1} can be recasted in the form of a Sternheimer-like equations for the excited states. In particular, the left action eigenvalue equation provides   
\begin{align}\lb{ExcitationOperatorsDef3}
&\left[E-(\hnot - \eps_p)\right] \ket{\phi_p^E} = \op Q_0 \op V'[\op E] \ket{\psi_p} \;, \nn\\
&\left[-E-(\hnot - \eps_p)\right]\ket{\tilde\chi_p^E} = \op Q_0\op V'[\op E^t]\ket{\psi_p}  \;, 
\end{align}
and the orthonormalization condition \eqref{orthoExcitatioOpDef1} implies %among the excitation operators is expressed as 
\be\lb{ExcitedStateOrthNormDef1}
\sum_p \left(\brket{\tilde\phi_p^E}{\phi_p^{E'}} + \brket{\tilde\chi_p^{E'}}{\chi_p^E}\right) = \delta(E-E') \;. 
\ee
Equations \eqref{ExcitationOperatorsDef3} allows us to present a localization argument for the excited states using the same procedure adopted for fluctuation states in 
section \ref{FluctuationState}. Pursing this strategy we present a formal solution of \eqref{ExcitationOperatorsDef3} as follows 
\begin{align}
&\ket{\phi^E_p} = \op G_H(E+\eps_p)\ket{s_{\phi}} \;, \nn \\
&\ket{\tilde\chi^E_p} = \op G_H(-E+\eps_p)\ket{s_{\tilde\chi}} \;,
\end{align}
where $\op G_H$ is the Helmholtz Green function, while the (self-consistent) source term $s_\phi$ and $s_{\tilde\chi}$ collect the sum of the r.h.s of \eqref{ExcitationOperatorsDef3} 
plus the action of the KS potential on the excited states.

@@

The typical structure of the excitations landscape is depicted in figure \ref{ExcitationLandscape}. The plot evidences that below the first ionization energy $|\eps_h|$ only localized excitations are 
present. Conversely, in the intermediate energy interval that extends from $|\eps_h|$ to the absolute value of the deepest occupied orbital, the elements belonging to the discrete sector are embedded 
in a continuum of delocalized excitations. Lastly, for energies higher than $|\eps_1|$ only delocalized excitations are present.

The introduction of the excitation operators allows us to express the susceptibility functional \eqref{SusceptibilityDef1} through the spectral decomposition of the resolvent of the 
Liouvillian. Pursuing this procedure provides:
%The knowledge of the excitations spectrum allows us to expand the LR susceptibility in terms of excitations. Indeed we can make usage of the spectral decomposition of the resolvent
%of the Liouvillian. This procedure allows us to formally solve the Liouville equation \eqref{LiouvillianRhopomegaDef1} and we have 
\be\lb{RhopExcitationDef1}
\sop{{\cal R}}(\omega) \cdot   =
\left(\sum_E + \int\dd E   \right) \op E \,
\frac{\trace{\commutator{\dmnot}{\op{\tilde E}}\cdot}}{\omega-E} \;,
\ee 
where we have divided the two sectors in the sum.

>>>>

We restrict our analysis assuming that a set of hypothesis commonly realized in DFT holds (symmetry of the transition basis, specify better? see appendix?). Equations 
\eqref{ExcitationOperatorsDef1} can be recasted in the form of a Sternheimer-like equations for the excited states. In particular, the left action eigenvalues equations provide   
\begin{align}\lb{ExcitationOperatorsDef3}
&\left[E-(\hnot - \eps_p)\right] \ket{\phi_p^E} = \op Q_0 \op V'[\op E] \ket{\psi_p} \;, \nn\\
&\left[-E-(\hnot - \eps_p)\right]\ket{\tilde\chi_p^E} = \op Q_0\op V'[\op E^t]\ket{\psi_p}  \;, 
\end{align}
and the orthonormalization condition \eqref{orthoExcitatioOpDef1} implies %among the excitation operators is expressed as 
\be\lb{ExcitedStateOrthNormDef1}
\sum_p \left(\brket{\tilde\phi_p^E}{\phi_p^{E'}} + \brket{\tilde\chi_p^{E'}}{\chi_p^E}\right) = \delta(E-E') \;. 
\ee
Equations \eqref{ExcitationOperatorsDef3} allows us to present a localization argument for the excited states using the same procedure adopted for fluctuation states in 
section \ref{FluctuationState}. Pursing this strategy we present a formal solution of \eqref{ExcitationOperatorsDef3} as follows 
\begin{align}
&\ket{\phi^E_p} = \op G_H(E+\eps_p)\ket{s_{\phi}} \;, \nn \\
&\ket{\tilde\chi^E_p} = \op G_H(-E+\eps_p)\ket{s_{\tilde\chi}} \;,
\end{align}
where $\op G_H$ is the Helmholtz Green function, while the (self-consistent) source term $s_\phi$ and $s_{\tilde\chi}$ collect the sum of the r.h.s of \eqref{ExcitationOperatorsDef3} 
plus the action of the KS potential on the excited states.

@@

The typical structure of the excitations landscape is depicted in figure \ref{ExcitationLandscape}. The plot evidences that below the first ionization energy $|\eps_h|$ only localized excitations are 
present. Conversely, in the intermediate energy interval that extends from $|\eps_h|$ to the absolute value of the deepest occupied orbital, the elements belonging to the discrete sector are embedded 
in a continuum of delocalized excitations. Lastly, for energies higher than $|\eps_1|$ the excitations landscape takes contributions only from the delocalized sector.

The introduction of the excitation operators allows us to express the susceptibility functional \eqref{SusceptibilityDef1} through the spectral decomposition of the resolvent of the 
Liouvillian. Pursuing this procedure provides:
%The knowledge of the excitations spectrum allows us to expand the LR susceptibility in terms of excitations. Indeed we can make usage of the spectral decomposition of the resolvent
%of the Liouvillian. This procedure allows us to formally solve the Liouville equation \eqref{LiouvillianRhopomegaDef1} and we have 
\be\lb{RhopExcitationDef1}
\sop{{\cal R}}(\omega) \cdot   =
\left(\sum_E + \int\dd E   \right) \op E \,
\frac{\trace{\commutator{\dmnot}{\op{\tilde E}}\cdot}}{\omega-E} \;,
\ee 
%where we have divided the two sectors in the sum. 
where we have explicitly separated the contributions of the localized and continuous excitations in the spectral decomposition \eqref{RhopExcitationDef1}. 
The elements of the first sector correspond to the \emph{true poles} of the susceptibility while the others can be interpreted only in terms of the spectral function.
